\section{Going further}\label{going-further}

The lists below do not pretend to be not exhaustive but may be a good starting point for those who whish to dive deeper in the topic of pointcloud processing with Python.

\subsection{Software}

Libraries (in alphabetical order):

\begin{itemize}
\item \textbf{CGAL}, an open-source library for efficient and reliable geometric algorithms (in C++, with Python bidings)
\item \textbf{CloudComPy}, a Python wrapper for CloudCompare (see below)
\item \textbf{Open3D}, an open-source library for 3D data processing (in C++ and Python, with a 3D viewer app)
\item \textbf{PCL}, a standalone, large scale, open project for 2D/3D image and pointcloud processing (in C++, with Python bidings)
\item \textbf{PDAL}, an open-source library for translating and manipulating pointcloud data (in C++, with Python support)
\item \textbf{PyMeshLab}, a Python library that interfaces to MeshLab (see below)
\item \textbf{PyVista}, a library providing a pythonic interface to VTK (see below)
\item \textbf{VTK}, an open-source software for manipulating and displaying scientific data (in C++, with wrappers in Python, Java and Tcl)
\end{itemize}

Applications (in alphabetical order):

\begin{itemize}
\item \textbf{Blender}, an open-source  3D computer graphics software that may be used to visualize and process pointclouds (with Python scripting capabilities)
\item \textbf{CloudCompare}, an open-source 3D pointcloud (and triangular mesh) processing software (with Python scripting capabilities through \textbf{CloudComPy})
\item \textbf{MeshLab}, an open-source 3D triangular meshes (and pointclouds) processing and editing software (with Python scripting capabilities through \textbf{PyMeshLab})
\item \textbf{ParaView}, an open-source visualization application (with Python scripting capabilities)
\end{itemize}

\subsection{Resources}

Books (in reverse chronological order):

\begin{itemize}
\item Poux, F. (2025). \textit{3D Data Science with Python}. O'Reilly Media.
\item Liu, S., Zhang, M., Kadam, P., \& Kuo, C. C. J. (2021). \textit{3D Point Cloud Analysis: Traditional, Deep Learning, and Explainable Machine Learning Methods}. Springer.
\item Vosselman, G., \& Maas, H. G. (2010). \textit{Airborne and terrestrial laser scanning}. Whittles Publishing.
\item Samet, H. (2006). \textit{Foundations of multidimensional and metric data structures}. Morgan Kaufmann.
\item Schneider, P., \& Eberly, D. H. (2002). \textit{Geometric tools for computer graphics}. Elsevier.
\item Goulette, F. (1999). \textit{Modélisation 3D automatique : outils de géométrie différentielle}. Presses des Mines.
\end{itemize}

Videos (in alphabetical order):

\begin{itemize}
\item CloudCompare playlist on Daniel Girardeau-Montaut YouTube channel \url{www.youtube.com/@danielgirardeau-montaut9044} (last accessed in October 2025)
\item Florent Poux YouTube channel: \url{www.youtube.com/@FlorentPoux} (last accessed in October 2025)
\end{itemize}
